Reproducible Research in R: A Tutorial on how to Do the Same Thing More Than Once

### 2021-11-19 manuscript.Rmd

One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
To lower the threshold, the R package `repro` introduces helper functions that wrap and hide the complexity of the necessary more complicated (and powerful) tools.
The provided wrapper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
The `repro` package extends the [`usethis`](https://usethis.r-lib.org) [@usethis] package.
While the `usethis` package aims to simplifys the development of R packages, `repro` contains reproducibility-specific utilities.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These wrappers are merely an assistance system, and as users get more and more comfortable, they can use the underlying tools directly to solve more complex problems.

This tutorial assumes that the user is predominantly working in R, with the help of RStudio.
It describes basic steps that we expect to be representative of a small-scale psychological research project that relies on no external software or multistage data processing (for those requirements see section [Advanced Features]). 
Of course, your specific situation might involve additional, more specialized steps. 
After completing the tutorial, you should be able to customize your workflow accordingly.

@@

The the check functions represent an example of the assistance that `repro` provides is .
They verify that you have indeed installed and set up the required pieces of software for this workflow (here, we use the double-colon operator to access functions from the `repro` namespace; we chose to do so to clarify which functions are native `repro` functions):

@@

@@

Now that your system is set up for this tutorial, we will introduce you to version control with `Git`.
`Git` tracks files within one project folder.
We start by creating such a project folder with RStudio by clicking the menu item:

This creates a project containing a sample analysis, but you may turn any other template or existing R project into a reproducible research project using the functionality of `repro`.

`Git` tracks any files that you manually add to the repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy-sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can add something to the `.gitignore` file directly or with this command:

@@

This interplay between `Docker` and `Make` is relatively complicated because it resembles a chicken or egg problem.
We have computational steps (`Make`) that depend on the software environment (`Docker`) for which we again have computational steps that create it.
Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an `R Markdown` or require other software than R packages.

>>>>

One impediment to the widespread adoption of a standard for reproducible research is that learning to use the required tools can be quite time-intensive.
To lower the threshold, the R package `repro` introduces helper functions that simplify the use of complicated and powerful tools.
The `repro` package follows the format of the [`usethis`](https://usethis.r-lib.org) [@usethis] package, which provides helper functions to simplify the development of R packages.
The `repro` package provides similar helper functions, but focuses on reproducibility-specific utilities.
These helper functions guide end-users in the use of reproducibility tools, provide feedback about what the computer is doing and suggest what the user should do next.
We hope this makes reproducibility tools more accessible by enabling beginner-level users to detect their system's state accurately and act correspondingly [@parasuramanAutomationHumanPerformance2018, Chapter 8: "Automation and Situation Awareness"].
These wrappers are merely a support system; as users learn to use the underlying tools, they can rely less on `repro` and use these ttools directly to solve more complex problems.

This tutorial assumes that the user will be working predominantly in R, with the help of RStudio.
It describes basic steps that we expect to be relevant for small-scale psychological research projects that do not rely on external software or multistage data processing (for those requirements see section [Advanced Features]). 
Of course, your specific situation might involve additional, more specialized steps. 
After completing the tutorial, you should be able to customize your workflow accordingly.

@@

To verify that you have indeed installed and set up the required software for this workflow,
you can use the "check functions".
These also illlustrate how `repro` assists the user in setting up a reproducible workflow.
In the example below, we use the double-colon operator to explicitly indicate which functions originate in the `repro` package.
If the package is loaded (using `library("repro")`), it is not necessary to use this double-colon notation.

@@

## Creating an RStudio Project

We start by creating a project folder with RStudio by clicking the menu item:

This creates a project with a sample analysis.
This sample analysis consists of a single R Markdown document and a single data file.
The only special thing about the R Markdown document is the `repro` metadata that we will learn later about. 
However, you may turn any other template or existing R project into a reproducible research project by adding those `repro` metadata there.

@@

Now that your project is set up, we will introduce you to version control with `Git`.
`Git` does not automatically track all files in your project folder;
rather, you must manually add files to the `Git` repository.
To make sure you do not accidentally add files that you do not wish to share (e.g., privacy-sensitive data),
you can list specific files that you do not want to track in the `.gitignore` file.
You can also block specific filetypes; for example, to prevent accidentally sharing raw data.
You can add something to the `.gitignore` file directly or with this command:

@@

The interplay between `Docker` and `Make` resembles a chicken or egg problem.
We have computational steps (`Make`) that depend on the software environment (`Docker`) for which we again have computational steps that create it.
Users only require a deeper understanding of this interdependence when they either want to have more complex computational recipes than rendering an `R Markdown` or require other software than R packages.
